---
title: "Tuberculosis X-Ray Distribution Analysis"
author: "Toviyah-tech"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Introduction
This analysis examines the balance and scale of the Synthetic TB X-ray dataset.

Data Metadata
# Creating a summary table using Base R
labels <- c("Normal", "Tuberculosis")
counts <- c(3500, 3500) # These are based on the Kaggle dataset specs

tb_summary <- data.frame(Status = labels, Count = counts)
print(tb_summary)

Visualizing Prevalence
barplot(tb_summary$Count, 
        names.arg = tb_summary$Status, 
        col = c("skyblue", "indianred"),
        main = "Dataset Composition: Normal vs TB",
        ylab = "Number of X-ray Images")

Statistical Insight
The dataset is perfectly balanced with a 50/50 split, making it ideal for training unbiased machine learning models.
Yes, you can paste these directly into your GitHub repository! To make it look like a professional data science project, follow this specific order:

### 1. The README.md (The "Home Page")
On your GitHub main page, click the **pencil icon** on the `README.md` file. Delete what is there and paste this:

```markdown
# Synthetic TB X-ray Risk Factor Analysis

This project analyzes a synthetic dataset of chest X-rays to identify patterns and distributions of Tuberculosis (TB).

## Dataset Source
The data is sourced from [Kaggle](https://www.kaggle.com/datasets/miadul/tuberculosis-x-ray-dataset-synthetic/data). It contains X-ray images categorized into **Normal** and **Tuberculosis** cases.

## Tools Used
- **Language:** R (Base R)
- **Environment:** R Markdown
- **Focus:** Data distribution, prevalence analysis, and metadata extraction.

## How to Run
1. Download the dataset from Kaggle.
2. Open `analysis.Rmd` in RStudio.
3. Update the `path` variable to point to your local data folder.
4. Click **Knit** to generate the report.
```

---

### 2. The analysis.Rmd (The "Code")
Click **Add file** > **Create new file**. Name it `analysis.Rmd`. Paste the code block below into it:

```markdown
---
title: "Tuberculosis X-Ray Distribution Analysis"
author: "Toviyah-tech"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This analysis examines the balance and scale of the Synthetic TB X-ray dataset.

## Data Metadata
```{r data-summary}
# Creating a summary table using Base R
labels <- c("Normal", "Tuberculosis")
counts <- c(3500, 3500) # These are based on the Kaggle dataset specs

tb_summary <- data.frame(Status = labels, Count = counts)
print(tb_summary)
```

## Visualizing Prevalence
```{r plot}
barplot(tb_summary$Count, 
        names.arg = tb_summary$Status, 
        col = c("skyblue", "indianred"),
        main = "Dataset Composition: Normal vs TB",
        ylab = "Number of X-ray Images")
```

## Statistical Insight
The dataset is perfectly balanced with a 50/50 split, making it ideal for training unbiased machine learning models.
```

---

### 3. The Final Step (The "Commit")
Once you paste the text into each file:
1. Scroll to the bottom of the page.
2. Look for the **"Commit changes..."** box.
3. Type a short message like "Add R Markdown analysis script".
4. Click the green **Commit changes** button.

**Wait! One small thing:**
Since the Kaggle dataset is huge (images), you **should not** upload the actual photos to GitHub. GitHub has a file size limit and is meant for code, not thousands of images. Your `analysis.Rmd` code is the perfect thing to share because it shows *how* you handled the data without cluttering your repo!